---
title: "Analysis of Historical Economic Data"
author: "Shivangi"
output: github_document
---

For this project I used a dataset of historical economic data derived from http://www.jerrydwyer.com/growth/index.html.
The data variables are described below and I'll be investigating which variables can be used to predict `ypp`, the output per-capita.


| Variable   | Description                                       |
| :--------- | :------------------------------------------------ |
| country    | Country Name                                      |
| region     | Region name                                       |
| year       | Year                                              |
| ypp        | Output Per-Capita                                 |
| school     | Years of Schooling (Avg.)                         |
| avgage     | Average Age (not Enrolled Under 65)               |
| exper      | Average Experience of Labor Force Participant     |
| hum_cap    | Level of Human Capital (Schooling and Experience) |
| lf2        | Labor Force                                       |
| pop        | Population                                        |
| sscap      | Steady State Capital Stock                        |
| kpw        | Capital Per Worker (Perpetual Inventory)          |
| lfpr       | Labor Force Particpation Rate                     |


```{r}
suppressPackageStartupMessages(library(tidyverse))

econ <- 
  read_csv("historical_econ.csv", show_col_types = FALSE)

econ_updated <-
  econ %>% 
  select(-sscap, -kpw, -lf2) %>% 
  pivot_longer(cols = c("ypp", "school", "avgage", "exper", "hum_cap"), names_to = "names", values_to = "values")

econ_joined <-
  econ_updated %>% 
  full_join(econ_updated, by = c("country", "region", "year", "pop", "lfpr")) %>% 
  filter(names.x != names.y)

ggplot(econ_joined) +
  geom_point(aes(x = values.x, y = values.y, color = year), size = 0.25, alpha = 0.3) +
  facet_grid(names.y ~ names.x, scales = "free")
```

This plot shows the relationship between each Variable with every other variable. We can also see how this relationship changes with time.
We can notice a very clear linear relationship between level of human capital and Average years of schooling. Here, Average years of schooling and Level of Human Capital increases over time. 

```{r}
suppressPackageStartupMessages(library(leaps))

forwardSelection <-
  leaps::regsubsets(
    ypp ~ . -country -region -year,
    method = "forward",
    data = econ,
    nvmax = 10) %>% 
  summary()

forwardSelection$which
index <- which.min(forwardSelection$cp)
colnames(forwardSelection$which)[forwardSelection$which[index,]]
forwardSelection$cp

fullModel <- lm(ypp ~ 1 -country -region -year, econ)
summary(fullModel)

bestModel <- lm(ypp ~ avgage + exper + hum_cap + kpw, econ)
summary(bestModel)
```

```{r}
suppressPackageStartupMessages(library(leaps))

backwardSelection <-
  leaps::regsubsets(
    ypp ~ . -country -region -year,
    method = "backward",
    data = econ,
    nvmax = 10) %>% 
  summary()

backwardSelection$which
index <- which.min(backwardSelection$cp)
colnames(backwardSelection$which)[backwardSelection$which[index,]]
backwardSelection$cp

fullModel <- lm(ypp ~ 1 -country -region -year, econ)
summary(fullModel)

bestModel <- lm(ypp ~ avgage + exper + hum_cap + kpw, econ)
summary(bestModel)
```

Selected predictors are not same each number of predictor.

```{r}
tibble(names = names(backwardSelection$which[2,]),
       forward = forwardSelection$which[2,],
       backward = backwardSelection$which[2,])

table(forwardSelection$which == backwardSelection$which)
```

```{r}
suppressPackageStartupMessages(library(modelr))
kcv <- 
  crossv_loo(econ) %>% 
  mutate(mod = map(train, 
                   ~lm(ypp ~ avgage + exper + hum_cap + kpw, data = .)
                   )) %>% 
  mutate(mse = map2_dbl(mod, test, mse))

mean(kcv$mse)/1000000
(mean(kcv$mse) + 1.96*sd(kcv$mse)/sqrt(nrow(econ)))/1000000
(mean(kcv$mse) - 1.96*sd(kcv$mse)/sqrt(nrow(econ)))/1000000
```