---
title: "MA 615 Homework 8"
author: "Shivangi"
output: github_document
---

For this HW we will use a dataset of historical economic data derived from http://www.jerrydwyer.com/growth/index.html.
The data variables are described below and we'll be investigating which variables can be used to predict `ypp`, the output per-capita.


| Variable   | Description                                       |
| :--------- | :------------------------------------------------ |
| country    | Country Name                                      |
| region     | Region name                                       |
| year       | Year                                              |
| ypp        | Output Per-Capita                                 |
| school     | Years of Schooling (Avg.)                         |
| avgage     | Average Age (not Enrolled Under 65)               |
| exper      | Average Experience of Labor Force Participant     |
| hum_cap    | Level of Human Capital (Schooling and Experience) |
| lf2        | Labor Force                                       |
| pop        | Population                                        |
| sscap      | Steady State Capital Stock                        |
| kpw        | Capital Per Worker (Perpetual Inventory)          |
| lfpr       | Labor Force Particpation Rate                     |


## Question 1

```{r}
suppressPackageStartupMessages(library(tidyverse))

econ <- 
  read_csv("historical_econ.csv", show_col_types = FALSE)

econ_updated <-
  econ %>% 
  select(-sscap, -kpw, -lf2) %>% 
  pivot_longer(cols = c("ypp", "school", "avgage", "exper", "hum_cap"), names_to = "names", values_to = "values")

econ_joined <-
  econ_updated %>% 
  full_join(econ_updated, by = c("country", "region", "year", "pop", "lfpr")) %>% 
  filter(names.x != names.y)

ggplot(econ_joined) +
  geom_point(aes(x = values.x, y = values.y, color = year), size = 0.25, alpha = 0.3) +
  facet_grid(names.y ~ names.x, scales = "free")
```

This plot shows the relationship between each Variable with every other variable. We can also see how this relationship changes with time.
We can notice a very clear linear relationship between level of human capital and Average years of schooling. Here, Average years of schooling and Level of Human Capital increases over time. 

## Question 2

```{r}
suppressPackageStartupMessages(library(leaps))

forwardSelection <-
  leaps::regsubsets(
    ypp ~ . -country -region -year,
    method = "forward",
    data = econ,
    nvmax = 10) %>% 
  summary()

forwardSelection$which
index <- which.min(forwardSelection$cp)
colnames(forwardSelection$which)[forwardSelection$which[index,]]
forwardSelection$cp

fullModel <- lm(ypp ~ 1 -country -region -year, econ)
summary(fullModel)

bestModel <- lm(ypp ~ avgage + exper + hum_cap + kpw, econ)
summary(bestModel)
```

## Question 3

```{r}
suppressPackageStartupMessages(library(leaps))

backwardSelection <-
  leaps::regsubsets(
    ypp ~ . -country -region -year,
    method = "backward",
    data = econ,
    nvmax = 10) %>% 
  summary()

backwardSelection$which
index <- which.min(backwardSelection$cp)
colnames(backwardSelection$which)[backwardSelection$which[index,]]
backwardSelection$cp

fullModel <- lm(ypp ~ 1 -country -region -year, econ)
summary(fullModel)

bestModel <- lm(ypp ~ avgage + exper + hum_cap + kpw, econ)
summary(bestModel)
```

## Question 4

Is the best model (according to `cp`) the same in Q2 and Q3?

That is true, the best model in question 2 and 3 is same

Are all of the selected predictors for backwards and forwards selection the same for each number of predictors? If not which are different?

Selected predictors are not same each number of predictor.

```{r}
tibble(names = names(backwardSelection$which[2,]),
       forward = forwardSelection$which[2,],
       backward = backwardSelection$which[2,])

table(forwardSelection$which == backwardSelection$which)
```


If you were to choose a model using just three predictors, which would you choose?

I will choose the one with lowest cp. Backward Selection Model in this case.

```{r}
forwardSelection$cp[3]
backwardSelection$cp[3]
```


```{r}
suppressPackageStartupMessages(library(modelr))
kcv <- 
  crossv_loo(econ) %>% 
  mutate(mod = map(train, 
                   ~lm(ypp ~ avgage + exper + hum_cap + kpw, data = .)
                   )) %>% 
  mutate(mse = map2_dbl(mod, test, mse))

mean(kcv$mse)/1000000
(mean(kcv$mse) + 1.96*sd(kcv$mse)/sqrt(nrow(econ)))/1000000
(mean(kcv$mse) - 1.96*sd(kcv$mse)/sqrt(nrow(econ)))/1000000
```